---
title: "Computational Modeling - Week 5 - Assignment 2 - Part 2"
author: "Riccardo Fusaroli"
date: "2/19/2017"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci.

### Second part: Focusing on predictions

Last year you assessed the teachers (darned time runs quick!). Now you want to re-test them and assess whether your models are producing reliable predictions. In Methods 3 we learned how to do machine-learning style assessment of predictions (e.g. rmse on testing datasets). Bayesian stats makes things a bit more complicated. So we'll try out how that works. N.B. You can choose which prior to use for the analysis of last year's data.

Questions to be answered (but see guidance below):
1- Write a paragraph discussing how assessment of prediction performance is different in Bayesian vs. frequentist models


2- Provide at least one plot and one written line discussing prediction errors for each of the teachers.

This is the old data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Josh: 160 correct answers out of 198 questions (Josh never gets bored)
- Mikkel: 66 correct answers out of 132 questions

This is the new data:
- Riccardo: 9 correct answers out of 10 questions (then he freaks out about teaching preparation and leaves)
- Kristian: 8 correct answers out of 12 questions
- Josh: 148 correct answers out of 172 questions (again, Josh never gets bored)
- Mikkel: 34 correct answers out of 65 questions

Guidance Tips

1. There are at least two ways of assessing predictions.
2. Last year's results are this year's expectations. We use the information from last year to make predictions for this year, as the prior. So, we will see whether our new data will be 'confirmed' by the prior (our posterior from last year)
3. Are the parameter estimates changing? (way 1)
4. How does the new data look in last year's predictive posterior? (way 2)


```{r}
setwd("~/Desktop/OneDrive/Cognitive Science/4 semester/Computational Models")

library(pacman)
p_load(rethinking, tidyverse)

```

## 1 Write a paragraph discussing how assessment of prediction performance is different in Bayesian vs. frequentist models

In frequentist models we calculate a specific value such as RMSE which we use as measure of prediction error. However, the RMSE value tells us nothing about the uncertainty inherent in the value, so there is no easy way to tell how probable the value is. 
In Bayesian assesment of prediction performance however, the output is a distribution of likely errors. It is much easier to see the uncertainty inherent in the model. In frequentist models we make predictions of the model using only data from the current experiment, where in bayesian models, we encode past knowledge of similar experiments asa prior, which we combine with the current experiment data to make a conclusion. So, in bayesian statistics we acknowledge that there is a prior probability. 

## 2 Provide at least one plot and one written line discussing prediction errors for each of the teachers

#first; create a function to calculate posterior given number of hits, n of possibilities and size of grid
```{r}

post = function(density, p_grid, right, questions, prior){
#compute likelihood at each value in grid
likelihood  <- dbinom(right, size = questions, prob=p_grid)
#compute product of likelihood and prior
unstd.posterior <- likelihood * prior
#standardize the posterior, so it sums to 1
posterior <- unstd.posterior / sum(unstd.posterior)

#return values in a dataframe
return(Data = data.frame(grid = p_grid, posterior=posterior, likelihood = likelihood, prior=prior))
}

#define density and pgrid and prior
dense = 1000
#define grid
p_grid <- seq(0,1, length.out = dense)
flatPrior <- rep(1,dense) #flat prior

# defining posteriors from old data portfolio 1 
riccardo = post(dense, p_grid, 3, 6, flatPrior)
josh = post(dense, p_grid, 160, 198, flatPrior)
mikkel = post(dense, p_grid, 66, 132, flatPrior)
kristian = post(dense, p_grid, 2, 2, flatPrior)

#RIC: 

# sample function 
samplez <- function(df){
  sample(df$grid, prob = df$posterior, size=1e5, replace=TRUE)
}
# see old ric. 
oldric = samplez(riccardo)
dens(oldric)

#creating new ric with the old ric as posterior
newric <- rbinom (1e4, size =10, prob=oldric)
#plot new ric
simplehist(newric)
dens(newric)


PI(newric,0.9) 

# to compare newric values with oldric, we compare by looking at the plot and seeing how much we need to adjust PI to make it fit. So to see when Riccardos number of correct questions, which is 9, we change PI between 0,1-1. It needs to be 0,9 before 9 is included in the prediction. Later on, we plot both the oldric and newric models in same plot to illustrate the difference. 

```

90 % of the probability distribution should be included in the intervals before riccardo's correct answer is included. Therefore, the model does not accurately predict performance. 


#Josh: 

```{r}
# sample function 
samplez <- function(df){
  sample(df$grid, prob = df$posterior, size=1e5, replace=TRUE)
}
# see old ric. 
oldjosh = samplez(josh)
dens(oldjosh)

#creating new ric with the old ric as posterior
newjosh <- rbinom (1e4, size =172, prob=oldjosh)
#plot new ric
simplehist(newjosh)
dens(newjosh)


HPDI(newjosh,0.9) #bad
```
 90 % of the probability values (probability mass) should be included in the intervals before josh's correct answer is included. Therefore, the model does not accurately predict performance. 


## Kristian
```{r}

# sample function 
samplez <- function(df){
  sample(df$grid, prob = df$posterior, size=1e5, replace=TRUE)
}
# see old ric. 
oldkris = samplez(kristian)
dens(oldkris)

#creating new ric with the old ric as posterior
newkris <- rbinom (1e4, size =12, prob=oldkris)
#plot new ric
simplehist(newkris)
dens(newkris)

# to see when kristians correct answers is included. 
HPDI(newkris,prob =0.7) 


```

70 % of the probable values (probability mass) should be included in the intervals before kristian's correct answer is included. Therefore, the model does not accurately predict performance. 


## Mikkel
```{r}

# sample function 
samplez <- function(df){
  sample(df$grid, prob = df$posterior, size=1e5, replace=TRUE)
}
# see old ric. 
oldmik = samplez(mikkel)
dens(oldmik)

#creating new ric with the old ric as posterior
newmik <- rbinom (1e4, size =65, prob=oldmik)
#plot new ric
simplehist(newmik)
dens(newmik)

# to see when Mikkels correct answer is included
PI(newmik,0.2) #pretty good prediction


```

20 % of the probability mass should be included in the intervals before mikkels's correct answer is included. Therefore, the model accurately predict performance. you go mikkel


```{r}
# compare old models with the new models to see whether the prediction fits. so the red line is the old model, the black is the new ones. 

newnewric <- post(dense, p_grid, 9, 10, riccardo$posterior)
dens(newnewric)

ggplot(newnewric,aes(grid,posterior))+ 
  geom_point()+
  geom_line()+
  theme_classic()+ 
  geom_line(aes(grid,prior,color='red'))+  xlab("cog sci knowledge")+ 
  ylab("posterior probability")

# pretty damn bad

newnewmik <- post(dense, p_grid, 34, 65, mikkel$posterior)
dens(newnewric)

ggplot(newnewmik,aes(grid,posterior))+ 
  geom_point()+
  geom_line()+
  theme_classic()+ 
  geom_line(aes(grid,prior,color='red'))+  xlab("cog sci knowledge")+ 
  ylab("posterior probability")

# really really good, u go mikkel

newnewjosh <- post(dense, p_grid, 34, 65, josh$posterior)

ggplot(newnewjosh,aes(grid,posterior))+ 
  geom_point()+
  geom_line()+
  theme_classic()+ 
  geom_line(aes(grid,prior,color='red'))+  xlab("cog sci knowledge")+ 
  ylab("posterior probability")

# okidoki, its good 

newnewkris <- post(dense, p_grid, 34, 65, kristian$posterior)

ggplot(newnewkris,aes(grid,posterior))+ 
  geom_point()+
  geom_line()+
  theme_classic()+ 
  geom_line(aes(grid,prior,color='red'))+  xlab("cog sci knowledge")+ 
  ylab("posterior probability")

# u not good kristian 
```


### Depending on time: Questions from the handbook
2H1. Suppose there are two species of panda bear. Both are equally common in the wild and live in the same places. They look exactly alike and eat the same food, and there is yet no genetic assay capable of telling them apart. They differ however in their family sizes. Species A gives birth to twins 10% of the time, otherwise birthing a single infant. Species B births twins 20% of the time, otherwise birthing singleton infants. Assume these numbers are known with certainty, from many years of field research.
Now suppose you are managing a captive panda breeding program. You have a new female panda of unknown species, and she has just given birth to twins. What is the probability that her next birth will also be twins?

2H2. Recall all the facts from the problem above. Now compute the probability that the panda we have is from species A, assuming we have observed only the first birth and that it was twins.

2H3. Continuing on from the previous problem, suppose the same panda mother has a second birth and that it is not twins, but a singleton infant. Compute the posterior probability that this panda is species A.

2H4. A common boast of Bayesian statisticians is that Bayesian inference makes it easy to use all of the data, even if the data are of different types. So suppose now that a veterinarian comes along who has a new genetic test that she claims can identify the species of our mother panda. But the test, like all tests, is imperfect. This is the information you have about the test:
- The probability it correctly identifies a species A panda is 0.8.
- The probability it correctly identifies a species B panda is 0.65.
The vet administers the test to your panda and tells you that the test is positive for species A. First ignore your previous information from the births and compute the posterior probability that your panda is species A. Then redo your calculation, now using the birth data as well.
